\documentclass{report}
\usepackage{graphicx}
\usepackage{listings}
\usepackage{color}
\usepackage[margin=1.2in]{geometry}
\usepackage{courier}
\usepackage{float}
\usepackage[toc,page]{appendix}

\definecolor{codegreen}{rgb}{0,0.6,0}
\definecolor{codeblack}{rgb}{0,0,0}
\definecolor{codered}{rgb}{0.867,0,0}
\definecolor{backcolour}{rgb}{ 0.95,0.95,0.95}
\definecolor{codeorange}{rgb}{1,0.447059,0}
\definecolor{codegrey}{rgb}{0.1, 0.1, 0.1}
\definecolor{codeblue}{rgb}{0.29, 0.52, 0.56}
\lstdefinestyle{mystyle}{
    basicstyle=\ttfamily\color{codeblack},
    backgroundcolor=\color{backcolour},   
    commentstyle=\color{codered},
    keywordstyle=\color{codeorange},
    numberstyle=\color{codeblack},
    stringstyle=\color{codegreen},
    breakatwhitespace=false,         
    breaklines=true,                 
    captionpos=b,                    
    keepspaces=true,                 
    numbers=left,                    
    numbersep=10pt,                  
    showspaces=false,                
    showstringspaces=false,
    showtabs=false,                  
    tabsize=2
}
 
\lstset{style=mystyle}

\newcommand{\ul}[1]{\underline{#1}}
\newcommand{\ull}[1]{\underline{\underline{#1}}}
\newcommand{\triangleBrac}[1]{\textless{#1}\textgreater}
\newcommand{\un}{\textunderscore}

\begin{document}
\title{CPS2000  - Compiler Theory and Practice - Assignment Report}
\author{Daniel Cauchi}
\date{}
\maketitle

\tableofcontents

\chapter{Introduction}
This is the report for the CPS2000 Compilers assignment, where a compiler and interpreter for the MiniLang programming language was created, with a Lexer, Parser, XML generator, Semantic Analyser and Interpreter. The C++ programming language was used in conjunction with the VSCode IDE. The program was tested on Ubuntu C++11, using the provided CMakeLists.txt. It works without error and any compiler warnings were dealt with. A video showcase is also attached to show the compiler in action, with several input files to show different aspects of it working in good order.

\chapter{Tasks}
\input{SubFiles/Question1}
\pagebreak
\input{SubFiles/Question2}
\pagebreak
\input{SubFiles/Question3}
\pagebreak
\input{SubFiles/Question4}
\pagebreak
\input{SubFiles/Question5}
\pagebreak

\chapter{Conclusion}
This concludes the report for the CPS2000-Compilers Assignment. For further information view the attached readme to get info on directory structure and how to use the compiler. The video showcase includes some examples which show the compiler's functionality and features in action.

\pagebreak

\bibliography{main}
\bibliographystyle{ieeetr}

\end{document}